---
date: "2022-01-24"
title: "Class 2"
menu:
  example:
    parent: Examples
weight: 2
type: docs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, class.source="language-r", class.output="language-r", message = FALSE, warning = FALSE)
xaringanExtra::use_clipboard()
library(rethinking)
```


```{r echo=FALSE}
downloadthis::download_file(
  path = "02-class.Rmd",
  output_name = "02-class",
  button_label = "Download this code",
  button_type = "danger",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

## Introduction

For this class, we'll review code examples found in Chapter 2.

This assumes that you have already installed the `rethinking` package.

If you need help, be sure to remember the references in the [Resources](/resource/):

* [Installing R/RStudio](/resource/install/)
* [Installing `rethinking` package](/resource/install-rethinking/)
* [Rmarkdown](/resource/rmarkdown/)
* [R Style guide](/resource/style/)

### Bayesian Updating: Grid Approximation

Let's assume we have the table in 2.1

```{r}
## R code 2.1
ways <- c( 0 , 3 , 8 , 9 , 0 )
ways/sum(ways)
```

Let's compute the likelihood of six W's in nine tosses (assuming a 50% probability):

```{r}
## R code 2.2
dbinom( 6 , size=9 , prob=0.5 )
```

We can see it's 16.4%.

Be sure to examine the `dbinom` function by typing in `?dbinom` and exploring the documentation. We'll use this function a lot in this class.

Next, let's define a grid. This is required when we are using Grid Approximation for our Bayesian calculations (i.e., to estimate the posterior).

```{r}
## R code 2.3
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

p_grid
```

Notice that this function creates a vector with length 20 and that ranges from 0 to 1. Note as well that each vector element is evenly spaced in increments of `(to - from)/(length.out - 1)`.

Think about the trade-off between having a smaller or larger `length.out`.

Next, let's define our prior. We'll assume a "flat" prior. 

```{r}
# define prior
prior <- rep( 1 , 20 )

prior
```

```{r}
# compute likelihood at each value in grid
likelihood <- dbinom( 6 , size=9 , prob=p_grid )

likelihood
```

```{r}
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

unstd.posterior
```

```{r}
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

posterior
```

```{r}
## R code 2.4
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```
What happens if we alter the priors? What will be the new posteriors?

```{r}
## R code 2.5
prior <- ifelse( p_grid < 0.5 , 0 , 1 )
prior <- exp( -5*abs( p_grid - 0.5 ) )
```

### Bayesian Updating: Quadratic Approximation

```{r}
## R code 2.6
library(rethinking)
globe.qa <- quap(
    alist(
        W ~ dbinom( W+L ,p) ,  # binomial likelihood
        p ~ dunif(0,1)     # uniform prior
    ) ,
    data=list(W=6,L=3) )

# display summary of quadratic approximation
precis( globe.qa )
```

```{r}
## R code 2.7
# analytical calculation
W <- 6
L <- 3
curve( dbeta( x , W+1 , L+1 ) , from=0 , to=1 )
# quadratic approximation
curve( dnorm( x , 0.67 , 0.16 ) , lty=2 , add=TRUE )
```


```{r}
## R code 2.8
n_samples <- 1000
p <- rep( NA , n_samples )
p[1] <- 0.5
W <- 6
L <- 3
for ( i in 2:n_samples ) {
    p_new <- rnorm( 1 , p[i-1] , 0.1 )
    if ( p_new < 0 ) p_new <- abs( p_new )
    if ( p_new > 1 ) p_new <- 2 - p_new
    q0 <- dbinom( W , W+L , p[i-1] )
    q1 <- dbinom( W , W+L , p_new )
    p[i] <- ifelse( runif(1) < q1/q0 , p_new , p[i-1] )
}

## R code 2.9
dens( p , xlim=c(0,1) )
curve( dbeta( x , W+1 , L+1 ) , lty=2 , add=TRUE )
```

### `tidyverse` conversion

Statistical Rethinking uses base R functions. More recently, Soloman Kurz has created a [translation of the book's functions](https://bookdown.org/content/4857/) into `tidyverse` (and later `brms`) code. This is not necessary but could be extremely helpful to classmates who are familiar with `tidyverse` already.

First, we'll need to call `tidyverse`. If you do not have `tidyverse`, you'll need to install it.

```{r}
library(tidyverse)
```


For example, we can translate 2.3 code using pipes (`%>%`)

```{r}
d <- tibble(p_grid = seq(from = 0, to = 1, length.out = 20),      # define grid
           prior  = 1) %>%                                       # define prior
    mutate(likelihood = dbinom(6, size = 9, prob = p_grid)) %>%  # compute likelihood at each value in grid
    mutate(unstd_posterior = likelihood * prior) %>%             # compute product of likelihood and prior
    mutate(posterior = unstd_posterior / sum(unstd_posterior)) 

d
```
With this calculated, we can then use `ggplot2`, the staple `ggplot2` data visualization package, to plot our posterior.

```{r}
d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_point() +
  geom_line() +
  labs(subtitle = "20 points",
       x = "probability of water",
       y = "posterior probability") +
  theme(panel.grid = element_blank())
```

For this class, we'll occasionally refer to Soloman's guide. 

## Package versions

```{r}
sessionInfo()
```