---
date: "`r Sys.Date()`"
title: "Class 15"
---

```{r setup, include=FALSE, fig.width=5, fig.height=4}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## In-Class Lab: Chapter 16

In this lab, we'll go through two examples from Chapter 16 with models that go beyond generalized linear models.

## Geometric People

Let's review the `Howell1` dataset and scale the values per 16.1 to eliminate the units.

```{r}
library(rethinking)
data(Howell1)
d <- Howell1

# scale observed variables
d$w <- d$weight / mean(d$weight)
d$h <- d$height / mean(d$height)
```

We can follow the code in lecture to view both plots.

```{r}
plot( d$height , d$weight , xlim=c(0,max(d$height)) , ylim=c(0,max(d$weight)) , col=col.alpha(2,0.7) ,
    lwd=3 , xlab="height (cm)" , ylab="weight (kg)" )
mw <- mean(d$weight)
mh <- mean(d$height)
lines( c(mh,mh) , c(0,mw) , lty=3 , lwd=2 )
lines( c(0,mh) , c(mw,mw) , lty=3 , lwd=2 )

plot( d$h , d$w , xlim=c(0,max(d$h)) , ylim=c(0,max(d$w)) , col=col.alpha(2,0.7) ,
    lwd=3 , xlab="height (scaled)" , ylab="weight (scaled)" )
lines( c(1,1) , c(0,1) , lty=3 , lwd=2 )
lines( c(0,1) , c(1,1) , lty=3 , lwd=2 )
```

### Scientific Model

We'll start with the scientific model for the volume of a cylinder:

$V = \pi r^{2} h$

By following the algebra in the textbook, we can rearrange to get this equation:

$W = kV = k \pi p^{2} h^{3}$

where $k$ is a constant.

Notice that given we have a scientific model, we no longer need a DAG.

### Statistical Model

We can rewrite our scientific model as:

$W_{i} ~ \sim Lognormal(\mu_{i},\sigma)$

$exp(\mu_{i}) = k \pi p^{2} h^{3}$

$k \sim some\,\,prior$

$p \sim some\,\,prior$

$\sigma \sim Exponential(1)$

We selected the Lognormal as the response function since weight can't be negative and is continuous.

### Prior Predictive Check

The key will be to select the priors for $k$ and $p$. We'll select $p \sim Beta(25,50)$ given that this would imply the prior will have a mean of $25 / (25 + 50) = 1/3$, given we're selecting a $Beta$ distribution. For $k$, we'll use $Exponential(0.5)$ given that this value represents the kilograms per cubic centimeters. Note that the priors shown in Lecture 19 are slightly different than what's in the textbook. We'll use Lecture 19 values.

```{r}
n <- 30
p <- rbeta(1e4,25,50)
k <- rexp(1e4,0.5)
sigma <- rexp(n,1)
xseq <- seq(from=0,to=1.3,len=100)
plot(NULL,xlim=c(0,1.3),ylim=c(0,1.5),xlab="height (scaled)",ylab="weight (scaled)")
for ( i in 1:n ) {
    mu <- log( pi * k[i] * p[i]^2 * xseq^3 )
    lines( xseq , exp(mu + sigma[i]^2/2) , lwd=3 , col=col.alpha(2,runif(1,0.4,0.8)) )
}
lines( c(1,1) , c(0,1) , lty=3 , lwd=2 )
lines( c(0,1) , c(1,1) , lty=3 , lwd=2 )

curve( dbeta(x,25,50) , from=0, to=1 , xlim=c(0,1) , lwd=3 , col=2 , xlab="p")

curve( dexp(x,0.5) , from=0, to=5 , xlim=c(0,5) , lwd=3 , col=2 , xlab="k")
```

### Initial Model

Given our statistical model above, let's write out our model:

```{r}
dat <- list(W=d$w,H=d$h)
m16.1 <- ulam(
    alist(
        W ~ dlnorm( mu , sigma ),
        exp(mu) <- 3.141593 * k * p^2 * H^3,
        p ~ beta( 25 , 50 ),
        k ~ exponential( 0.5 ),
        sigma ~ exponential( 1 )
    ), data=dat , chains=4 , cores=4 )
```

### Fit Model

```{r}
precis(m16.1)
```

### Convergence

```{r}
traceplot(m16.1)
```

We find sufficient model convergence in the traceplots, consistent with the Rhat.

#### Posterior Predictive Simulation

Let's do a posterior predictive simulation.

```{r}
h_seq <- seq( from=0 , to=max(d$h) , length.out=100 )
w_sim <- sim( m16.1 , data=list(H=h_seq) )
mu_mean <- apply( w_sim , 2 , mean )
w_CI <- apply( w_sim , 2 , PI )
plot( d$h , d$w , xlim=c(0,max(d$h)) , ylim=c(0,max(d$w)) , col=2 ,
    lwd=2 , xlab="height (scaled)" , ylab="weight (scaled)" )
shade( w_CI , h_seq , col=col.alpha(2,0.5) )
lines( h_seq , mu_mean , lwd=3 )
```

We can see that with the exception of children (i.e., small weight/height), the cylinder model does a fairly good job of fitting the model. You can review Lecture 19 for more details on how we can rewrite different versions of this model.

## Population Dynamics

Let's also review the Lynx and Hare example in Lecture 19 and Chapter 16.

```{r}
## R code 16.13
library(rethinking)
data(Lynx_Hare)
Lynx_Hare
```

Let's plot the data.

```{r}

plot( 1:21 , Lynx_Hare[,3] , ylim=c(0,90) , xlab="year" ,
    ylab="thousands of pelts" , xaxt="n" , type="l" , lwd=3  )
at <- c(1,11,21)
axis( 1 , at=at , labels=Lynx_Hare$Year[at] )
lines( 1:21 , Lynx_Hare[,2] , lwd=3 , col=2 )
points( 1:21 , Lynx_Hare[,3] , bg="black" , col="white" , pch=21 , cex=1.4 )
points( 1:21 , Lynx_Hare[,2] , bg=2 , col="white" , pch=21 , cex=1.4 )
text( 17 , 80 , "Lepus" , pos=2 )
text( 19 , 50 , "Lynx" , pos=2 , col=2 )
```

### Generative Model

Let's create a generative model. This is based on the ordinary differential equations laid out in Lecture 19. See Lecture 19 for the intuition around the equations.

```{r}
## R code 16.14
sim_lynx_hare <- function( n_steps , init , theta , dt=0.002 ) {
    L <- rep(NA,n_steps)
    H <- rep(NA,n_steps)
    L[1] <- init[1]
    H[1] <- init[2]
    for ( i in 2:n_steps ) {
        H[i] <- H[i-1] + dt*H[i-1]*( theta[1] - theta[2]*L[i-1] )
        L[i] <- L[i-1] + dt*L[i-1]*( theta[3]*H[i-1] - theta[4] )
    }
    return( cbind(L,H) )
}

## R code 16.15
theta <- c( 0.5 , 0.05 , 0.025 , 0.5 )
z <- sim_lynx_hare( 1e4 , as.numeric(Lynx_Hare[1,2:3]) , theta )

plot( z[,2] , type="l" , ylim=c(0,max(z[,2])) , lwd=4 , xaxt="n" ,
    ylab="number (thousands)" , xlab="" )
lines( z[,1] , col=2 , lwd=4 )
mtext( "time" , 1 )
```

### Prior Predictive Simulation

```{r}
## R code 16.16
N <- 1e4
Ht <- 1e4
p <- rbeta(N,2,18)
h <- rbinom( N , size=Ht , prob=p )
h <- round( h/1000 , 2 )
dens( h , xlab="thousand of pelts" , lwd=2 )
```

### Initial Model

Let's specify our model.

```{r}
## R code 16.17
data(Lynx_Hare_model)
cat(Lynx_Hare_model) # Stan model specifications (see Lecture 19)
```
We can now run the model into `cstan`.

```{r}
## R code 16.18
dat_list <- list(
    N = nrow(Lynx_Hare),
    pelts = Lynx_Hare[,2:3] )

m16.5 <- cstan( model_code=Lynx_Hare_model , data=dat_list , chains=3 ,
    cores=3 , control=list( adapt_delta=0.95 ) )
```

```{r}
precis(m16.5, depth=2)
```

### Convergence

```{r}
traceplot(m16.5)
```

### Posterior Predictive Simulation

```{r}
## R code 16.19
post <- extract.samples(m16.5)
pelts <- dat_list$pelts
plot( 1:21 , pelts[,2] , pch=16 , ylim=c(0,140) , xlab="year" ,
    ylab="thousands of pelts" , xaxt="n" )
at <- c(1,11,21)
axis( 1 , at=at , labels=Lynx_Hare$Year[at] )
# 21 time series from posterior
for ( s in 1:21 ) {
    lines( 1:21 , post$pelts_pred[s,,2] , col=col.alpha("black",0.2) , lwd=3 )
    lines( 1:21 , post$pelts_pred[s,,1] , col=col.alpha(2,0.3) , lwd=3 )
}
# points
points( 1:21 , pelts[,2] , col="white" , pch=16 , cex=1.5 )
points( 1:21 , pelts[,2] , col=1 , pch=16 )
points( 1:21 , pelts[,1] , col="white" , pch=16 , cex=1.5 )
points( 1:21 , pelts[,1] , col=2 , pch=16 )
# text labels
text( 17 , 110 , "Lepus" , pos=2 )
text( 19 , 50 , "Lynx" , pos=2 , col=2 )
```



## Package versions

```{r}
sessionInfo()
```
